Cost- and Deadline-Constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds

### 2012-04-25 paper/paper.tex
Reduce conclusion more

% New rewritten text for section ``discussion of Results'' starts here

\begin{figure}[tb]
    \centering
    \subfloat[size]{
        \includegraphics[width=0.6\columnwidth]{size-selected-LIGO}
        }\\
    \subfloat[cost]{
        \includegraphics[width=0.6\columnwidth]{cost-selectedLIGO}
        }
    \caption{Example performance of algorithms expressed in
    sum of dag sizes and effective cost, data for LIGO application, Pareto
    Sorted ensemble of 100 workflows, budget = \$600}
    \label{fig:size-cost}
\end{figure}

%To analyze the behaviour of algorithms we ran simulations for constraint values
%ranging between the estimated limits indicated in
%Section~\ref{sec:exp-parameters}, resulting in $10 * 10 = 100$ (budget,
%deadline) data points for each of 5 applications, 3 algorithms and using 5
%ensemble size distributions and 10 random seeds. From this result space of
%75000 we selected representative plots.

To illustrate the behavior of the algorithms we show here examples of
performance curves obtained for selected sample parameters.
Fig.~\ref{fig:size-cost}a uses the total computing time metric, $T_C$ for LIGO
application with pareto Sorted distribution of workflows. It can be seen that
SPSS gives the best performance and WA-DPDS performs better than the
workflow-unaware version, DPDS. This means that for the workflows we tested the
simple admission procedure based on estimation of workflow cost does not degrade
the solution, but rather in many cases it enables larger workflows that would
lead to budget overrun to be rejected. Thus it can save the space for smaller
workflows to complete.

@@

The goal of the first experiment is to characterize the relative performance of the DPDS, WA-DPDS, and SPSS algorithms. This was done by simulating the algorithms on many different ensembles and comparing the scores computed using the technique described in Section~\ref{sec:perf_metric}. The algorithm that achieves the largest number of high scores is judged to have the best performance.

@@

In this paper we addressed the interesting and important new problem of scheduling and resource provisioning for scientific workflow ensembles on IaaS clouds. The problem is different from previous work on grid and utility grid scheduling in that cloud infrastructures provide more control over the resources, so the resource provisioning plan can be adjusted according to the requirements of the application. Therefore the problem space becomes larger; it requires not only an efficient mapping of tasks to available resources, but also the selection of the best resource provisioning plan. Formulating the problem as a maximization of the number of prioritized workflows completed from the ensemble is also novel and requires workflows to be admitted or rejected based on their estimated resource demands. We believe that this bi-constrained problem is highly relevant because such constraints are commonly imposed on many real-world projects. The approach is also directly applicable to grid environments that provide resource reservations and charge service units for resource use.

The DPDS, WA-DPDS and SPSS algorithms were developed based on both static and dynamic scheduling approaches. The algorithms were evaluated on ensembles of synthetic workflows, which were generated based on real scientific applications. For the purposes of evaluation we have developed a simulator that models the cloud infrastructure, and a workflow engine with tightly-coupled scheduler and provisioner modules.

The results of our simulation studies indicate that the two algorithms that take into account information about the structure of the workflow and task runtime estimates (WA-DPDS and SPSS) yield better results than the simple priority-based scheduling strategy (DPDS), which makes provisioning decisions purely based on resource utilization. On the other hand, DPDS may be the only choice when estimates of task runtimes are not available in advance.

In cases where there are no provisioning delays, task runtime estimates are good, and failures are rare, we found that the static algorithm performs significantly better than the dynamic algorithms. However, when runtime estimates are poor, delays are present, or failures are common, the plans produced by the static algorithm are obstructed and it frequently exceeds the budget and deadline constraints. In comparison, the dynamic algorithms are able to quickly adapt to even verly large delays, poor estimates, and failures to prevent the ensemble from exceeding the constraints.

In addition, we found that SPSS tends to perform better on coarse-grained workflows than the dynamic algorithms. For wide, fine-grained workflows, such as CyberShake and Montage, however, the dynamic algorithms frequently produce performance that is as good or better than SPSS,  because they are better able to pack the smaller tasks onto idle VMs close to the deadline than SPSS, which distributes deadlines to tasks in a way that prevents them from starting late.

For very large workflows and ensembles, we found that the planning time of the SPSS algorithm is prohibitive due to its $O(n^2)$ algorithmic complexity. Our simulation results showed that ensembles of 100 workflows with 800 tasks often took 10 minutes or more to plan. This suggests that ensembles of workflows with tens of thousands of tasks, which are commonly encountered in real workflow applications, would take many hours to plan. In comparison, the $O(n)$ complexity of the dynamic algorithms makes them much more attractive for large scale problems.

This study suggests many areas for future work. It would be interesting to extend the application and infrastructure model to explicitly include the various data storage options available on clouds (right now data access is modeled as part of the task execution time and data costs are not considered). A previous experimental study \cite{Juve2010} suggests that the data demands of scientific workflows have a large impact on not only the execution time, but also on the cost of workflows in commercial clouds. Finally, it would be interesting to investigate heterogeneous environments that include multiple VM types and cloud providers, including private and community clouds, which will make the problem even more complex and challenging~\cite{Marshall2010, Vockler2011, Juve2010}.

The results of this study can also be applied to develop tools that assist researchers in planning their large-scale computational experiments. The estimates of cost, runtime, and number of workflows completed that can be obtained from both the static algorithms and from the simulation runs, constitute valuable hints for planning ensembles and evaluating the associated trade-offs.

>>>>

\begin{figure}[tb]
    \centering
    \subfloat[size]{
        \includegraphics[width=0.6\columnwidth]{size-selected-LIGO}
    }\\
    \subfloat[cost]{
        \includegraphics[width=0.6\columnwidth]{cost-selectedLIGO}
    }
    \caption{Example performance of algorithms expressed in sum of workflow sizes and effective cost, data for LIGO application, Pareto Sorted ensemble of 100 workflows, budget = \$600}
    \label{fig:size-cost}
\end{figure}

%To analyze the behavior of the algorithms we ran simulations for constraint values ranging between the estimated limits indicated in Section~\ref{sec:exp-parameters}, resulting in $10 * 10 = 100$ (budget, deadline) data points for each of 5 applications, 3 algorithms and using 5 ensemble size distributions and 10 random seeds. From this result space of 75,000 we selected representative plots.

To illustrate the behavior of the algorithms we show here examples of performance curves obtained for selected sample parameters. Fig.~\ref{fig:size-cost}a uses the total computing time metric, $T_C$ for LIGO application with Pareto Sorted distribution of workflows. It can be seen that SPSS gives the best performance and WA-DPDS performs better than the workflow-unaware version, DPDS. This means that for the workflows we tested the simple admission procedure based on estimation of workflow cost does not degrade the solution, but rather in many cases it enables larger workflows that would lead to budget overrun to be rejected. Thus it can save the space for smaller workflows to complete.

@@

The goal of the the next experiment is to characterize the relative performance of the proposed algorithms. This was done by simulating the algorithms on many different ensembles and comparing the scores computed using the exponential scoring technique described in Section~\ref{sec:perf_metric}. The algorithm that achieves the largest number of high scores is judged to have the best performance.

@@

In this paper we addressed the interesting and important new problem of scheduling and resource provisioning for scientific workflow ensembles on IaaS clouds. The goal of this work is to maximize the number of prioritized workflows that can be completed given budget and deadline constraints.

%This problem differs from previous work on grid and utility grid scheduling in that cloud infrastructures provide more control over provisioning, so that the number of resources can be adjusted according to the requirements of the application. Therefore the problem space becomes larger; it requires not only an efficient mapping of tasks to available resources, but also the selection of the best resource provisioning plan.

%Formulating the problem as a maximization of the number of prioritized workflows completed from the ensemble is also novel and requires workflows to be admitted or rejected based on their estimated resource demands. We believe that this bi-constrained problem is highly relevant because such constraints are commonly imposed on many real-world projects. The approach is also directly applicable to grid environments that provide resource reservations and charge service units for resource use.

We developed three algorithms to solve this problem: two dynamic algorithms, DPDS and WA-DPDS, and one static algorithm, SPSS. The algorithms were evaluated via simulation on ensembles of synthetic workflows, which were generated based on statistics from real scientific applications.

The results of our simulation studies indicate that the two algorithms that take into account the structure of the workflow and task runtime estimates (WA-DPDS and SPSS) yield better results than the simple priority-based scheduling strategy (DPDS), which makes provisioning decisions based purely on resource utilization.

In cases where there are no provisioning delays, task runtime estimates are good, and failures are rare, we found that the static algorithm, SPSS, performs significantly better than both dynamic algorithms. However, when conditions are less than perfect, the static plans produced by SPSS are disrupted and it frequently exceeds the budget and deadline constraints. In comparison, the dynamic algorithms are able to adapt to a wide variety of conditions, and rarely exceed the constraints even with long delays, poor estimates, and high failure rates.

%In addition, we found that SPSS tends to perform better on coarse-grained workflows than the dynamic algorithms. For wide, fine-grained workflows, such as CyberShake and Montage, however, the dynamic algorithms frequently produce performance that is as good or better than SPSS, because they are better able to pack the smaller tasks onto idle VMs close to the deadline than SPSS, which distributes deadlines to tasks in a way that prevents them from starting late.

%For very large workflows and ensembles, we found that the planning time of the SPSS algorithm is prohibitive. SPSS often took 10 minutes or more to plan ensembles of 100 workflows with 800 tasks each. This suggests that ensembles of workflows with tens of thousands of tasks, which are commonly encountered in real workflow applications, would take many hours to plan using SPSS. This makes the dynamic algorithms much more attractive for large scale problems.

This study suggests several areas for future work. Our current approach models data access as part of the task execution time and does not consider data storage and transfer costs. In the future we plan to extend the application and infrastructure model to  include the various data storage options available on clouds. A previous experimental study \cite{Juve2010} suggests that the data demands of scientific workflows have a large impact on not only the execution time, but also on the cost of workflows in commercial clouds. We also plan to investigate heterogeneous environments that include multiple VM types and cloud providers, including private and community clouds, which will make the problem even more complex and challenging.

%The results of this study can be applied to develop tools that assist researchers in planning their large-scale computational experiments. The estimates of cost, runtime, and number of workflows completed that can be obtained from both the static algorithms and from the simulation runs, constitute valuable hints for planning ensembles and evaluating the associated trade-offs.
